Self-tunable approximated explicit MPC: Heat exchanger implementation and analysys

### 2023-02-03 paper/elsarticle/main.tex
Review: Sec.4

	The plant on which the control was implemented and analyzed is a laboratory-scaled counter-current liquid-to-liquid plate heat exchanger Armfield Process Plant Trainer PCT23~\cite{pct23}, see Figure~\ref{fig:HE}. The heat exchanger is 90 mm wide, 103 mm long, and 160 mm high. The heat exchange is performed between the cold feed (water) and the hot medium (water). The cold feed as well as the heating medium are transported to the heat exchanger by two peristaltic pumps with flexible tubing from silicon rubber. The flow rate of the feed is constant, while the aim of control is to track the reference value of the feed temperature. Therefore, the controlled variable is the feed temperature $T$ at the outlet of the heat exchanger. The temperature of the heated feed in the outlet stream was measured by the type K thermocouple. The associated manipulated variable is the voltage $U$ corresponding to the power of the pump feeding the heat exchanger by the hot medium. The voltage is within the range of [$0-5$]\,V normalized into the relative values in percentage. For further technical specifications of the laboratory heat exchanger, the reader is referred to~\cite{pct23}. As heat exchange is a nonlinear and asymmetric process~\cite{Liptak}, this heat exchanger represents a suitable candidate for the presented controller tuning strategy. The corresponding simplified closed-loop scheme can be seen in Figure~\ref{fig:CL}. Note, the ,,Self-tuning" block corresponds to the more detailed scheme in Figure~\ref{fig:tuning}.

@@

	The penalty matrices of the problem Eq.~\eqref{eq:mpc_problem} were systematically tuned and the corresponding control was implemented on the laboratory heat exchanger for each of the considered explicit MPC setups. 
	First, the aim of tuning was to determine, which penalty matrix is the most suitable for real-time tuning. Based on the set of experimentally collected data, the most significant effect on the control trajectories had tuning the penalty matrix $Q_\mathrm{y}$, while still preserving a satisfactory control performance, i.e., without steady-state control error and significant oscillations around the reference value. Next, the boundary values of the tunable matrix $Q_\mathrm{y}$ were tuned as $Q_\mathrm{y, L}$ = 100 and $Q_\mathrm{y, U}$ = 1000. The built-in integrator was penalized with the fixed penalty matrix $Q_\mathrm{x}$ = 1 and the control input with the fixed penalty matrix $R$ = 10. The prediction horizon $N$ was set to 20 control steps. The explicit model predictive controllers were constructed in MATLAB R2020b using the Multi-Parametric Toolbox 3~\cite{mpt_conf}.

@@

	\begin{figure}
		\begin{center}
			\includegraphics[width=\textwidth]{images/cl}
			\caption[Closed-loop scheme]{Closed-loop scheme.}
			\label{fig:CL}
		\end{center}
	\end{figure}

@@

	To analyze the control more objectively, energy consumption was also investigated. More specifically, the power used to heat the heating medium was measured and summarized for the whole control but for every controller setup, respectively. The energy consumption $E$ necessary for every control as well as the relative improvement subject to every boundary controller can be seen in Table~\ref{tab:energy}.

	\begin{table}[h!]
		\begin{center}
			\caption{Energy consumption evaluation.}
			\label{tab:energy}
			\begin{tabular}{c|c|c} 
				$Q_\mathrm{y}$ & E\,[kJ] & $\delta E\,[\%]$   \\
				\hline
				1000 &  883 & -4\\ 
				% \hline
				100 & 938 & 2 \\ 
				% \hline
				self-tuned & 921 & -   
			\end{tabular}
		\end{center}
	\end{table}

	It can be seen in Table~\ref{tab:energy}, that utilizing the self-tuned controller leads to energy consumption which acquires value between the energy consumptions associated with the boundary controllers. Although it is not the minimum one, the energy is still saved compared to the boundary controller with $Q_\mathrm{y} = 100$. Note, though the controller with $Q_\mathrm{y} = 1000$ leads to minimum energy consumption, using this single non-tuned controller does not lead to such control performance improvement.

	In general, utilizing the proposed controller with a scalable aggressiveness according to the operating conditions leads to higher accuracy (lower $ISE$), lower value of the overshoots (lower $\sigma_{\mathrm{max}}$), and faster achieving the reference value (lower $t_{\epsilon}$), while the control does not suffer in the terms of energy consumption.

>>>>

	The plant on which the control was implemented and analyzed is a laboratory-scaled counter-current liquid-to-liquid plate heat exchanger Armfield Process Plant Trainer PCT23~\cite{pct23}, see Figure~\ref{fig:HE}. The heat exchanger is 90\,mm wide, 103\,mm long, and 160\,mm high. The heat exchange is performed between the cold feed (water) and the hot medium (water). The cold feed as well as the heating medium are transported to the heat exchanger by two peristaltic pumps with flexible tubing from silicon rubber. The flow rate of the feed is constant, while the aim of control is to track the reference value of the feed temperature. Therefore, the controlled variable is the feed temperature $T$ at the outlet of the heat exchanger. The temperature of the heated feed in the outlet stream was measured by the type K thermocouple. The associated manipulated variable is the voltage $U$ corresponding to the power of the pump feeding the heat exchanger by the hot medium. The voltage is within the range of [$0-5$]\,V normalized into the relative values in percentage. For further technical specifications of the laboratory heat exchanger, the reader is referred to~\cite{pct23}. As heat exchange is a nonlinear and asymmetric process~\cite{Liptak}, this heat exchanger represents a suitable candidate for the presented controller tuning strategy. The corresponding illustrative scheme of the implemented closed-loop control setup is in Figure~\ref{fig:CL}, where the ``self-tuning'' block substitutes the more detailed scheme in Figure~\ref{fig:tuning}.

@@

	The penalty matrices of the problem Eq.~\eqref{eq:mpc_problem} were systematically tuned and the corresponding control was implemented on the laboratory heat exchanger for each of the considered explicit MPC setups. 
	First, the aim of tuning was to determine, which penalty matrix is the most suitable for real-time tuning. Based on the set of experimentally collected data, the most significant effect on the control trajectories had tuning the penalty matrix $Q_\mathrm{y}$, while still preserving a satisfactory control performance, i.e., without steady-state control error and significant oscillations around the reference value. Next, the boundary values of the tunable matrix $Q_\mathrm{y}$ were tuned as $Q_\mathrm{y, L}$ = 100 and $Q_\mathrm{y, U}$ = 1\,000. The built-in integrator was penalized with the fixed penalty matrix $Q_\mathrm{x}$ = 1 and the control input with the fixed penalty matrix $R$ = 10. The prediction horizon $N$ was set to 20 control steps. The explicit model predictive controllers were constructed in MATLAB R2020b using the Multi-Parametric Toolbox 3~\cite{mpt_conf}.

@@

	\begin{figure}
		\begin{center}
			\includegraphics[width=\textwidth]{images/cl}
			\caption[Scheme of the implemented closed-loop control setup.]{Closed-loop scheme.}
			\label{fig:CL}
		\end{center}
	\end{figure}

@@

	To analyze the control from the environmental viewpoint, the energy consumption was also investigated. Technically, the total power supply utilized to heat the heating medium was measured and summarized for the whole control but for each controller setup, respectively. The total energy consumption $E$ consumed by each controller as well as the relative improvement subject to each boundary controller is summarized in Table~\ref{tab:energy}.

	\begin{table}[h!]
		\begin{center}
			\caption{Evaluation of the total energy consumption.}
			\label{tab:energy}
			\begin{tabular}{c|c|c} 
				$Q_\mathrm{y}$ & E\,[kJ] & $\delta E\,[\%]$   \\
				\hline
				1000 &  883 & $-4$\\ 
				% \hline
				100 & 938 & $2$ \\ 
				% \hline
				self-tuned & 921 & $-$
			\end{tabular}
		\end{center}
	\end{table}

	It can be seen in Table~\ref{tab:energy}, that utilizing the self-tuned controller leads to energy consumption which acquires value between the energy consumption corresponding to the boundary controllers. Although it is not the minimum one, the energy is still saved compared to the boundary controller with $Q_\mathrm{y} = 100$. Note, though the controller with $Q_\mathrm{y} = 1\,000$ leads to minimum energy consumption, using this single non-tunable controller does not ensured the required control performance. As for the test bed served the laboratory-scaled heat exchanger plant, the greater energy savings are expected for the industrial-scaled plants.

	In general, utilizing the proposed controller with a scalable aggressiveness according to the operating conditions leads to higher accuracy (lower $ISE$), lower value of the overshoots (reduced $\sigma_{\mathrm{max}}$), and faster achieving the reference value (decreased $t_{\epsilon}$), while the control performance does not suffer in the terms of energy consumption.
